---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TidyTuesday data viewer

<!-- badges: start -->
<!-- badges: end -->

`ttdviewer` is an Shiny app-package[^1] for viewing the data from the [`#TidyTuesday`](https://github.com/rfordatascience/tidytuesday) project. This package started out as a small application to demonstrate various examples of downloadable reports (R Markdown, Quarto, etc.), but it's grown to include additional features: 

[^1]: An R package containing a Shiny application. Read more [here](https://mastering-shiny.org/scaling-packaging.html).

1. The data are loaded directly from GitHub (i.e., with help from the  [`tidytuesdayR`](https://dslc-io.github.io/tidytuesdayR/) and [`ttmeta`](https://r4ds.github.io/ttmeta/) packages). 
2. The returned object is visible as a list on the first `nav_panel()` via the [`listviewerlite` package](https://long39ng.github.io/listviewerlite/).   
3. The data are visible as a table compliments of the [`reactable` package](https://glin.github.io/reactable/) in the second `nav_panel()`.    
4. Data visualizations from the [`inspectdf` package](https://alastairrushworth.com/inspectdf/) '*summarise missingness, categorical levels, numeric distribution, correlation, column types and memory usage*' are available on the third `nav_panel()`.    


## Installation

You can install the development version of `ttdviewer` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mjfrigaard/ttdviewer")
```

## Launch app

Launch the application using the standalone app function below:

```{r}
#| label: example
#| eval: false
library(ttdviewer)
launch()
```
