---
title: "App"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: package
#| warning: false
#| message: false
library(dplyr)
library(tidyr)
library(inspectdf)
library(lobstr)
library(ttdviewer)
```

## App structure 

The application's functions are in the abstract syntax tree below:

```{r}
#| eval: false
#| include: false
ast(
  launch(
    app_ui(
      mod_input_ui(),
      mod_report_input_ui(),
      mod_report_desc_ui(),
      mod_report_download_ui(),
      mod_list_ui(),
      mod_plot_ui(),
      mod_table_ui()
    ),
    app_server(
      mod_input_server(),
      mod_report_input_server(),
      mod_report_desc_server(),
      mod_report_download_server(),
      mod_list_server(),
      mod_plot_server(),
      mod_table_server()
    )
  )
)
```

```{verbatim}
█─launch 
├─█─app_ui 
│ ├─█─mod_input_ui 
│ ├─█─mod_report_input_ui 
│ ├─█─mod_report_desc_ui 
│ ├─█─mod_report_download_ui 
│ ├─█─mod_list_ui 
│ ├─█─mod_plot_ui 
│ └─█─mod_table_ui 
└─█─app_server 
  ├─█─mod_input_server 
  ├─█─mod_report_input_server 
  ├─█─mod_report_desc_server 
  ├─█─mod_report_download_server 
  ├─█─mod_list_server 
  ├─█─mod_plot_server 
  └─█─mod_table_server
```

### Variable inputs 

The variable input (`input$dataset_title`) is collected in the `mod_input_ui()` and `mod_input_server()` functions and passed to the `load_tt_data()` function: 

```{r}
#| eval: false
#| code-fold: false
tryCatch({
    result <- load_tt_data(input$dataset_title)
    if (length(result) == 0) {
      logr_msg("Empty dataset returned", level = "WARN")
      showNotification("No data available for selected dataset",
        type = "warning", duration = 5
      )
    } else {
      logr_msg("Dataset successfully loaded in reactive",
        level = "SUCCESS"
      )
    }
    return(result)
  },
  error = function(e) {
    logr_msg(paste("Error in data reactive:", e$message),
      level = "ERROR"
    )
    showNotification(paste("Error loading data:", e$message),
      type = "error", duration = 10
    )
    return(list())
  }
)
```

`mod_input_server()` returns the dataset list and title as a reactive.
```{r}
#| eval: false
#| code-fold: false
return(
    list(
      data = data,
      'ds_title' = reactive(input$ds_title)
    )
  )
```

In the `app_server()`, the inputs are assigned to the `title_input` and separated into `selected_data` and `dataset_title`:

```{r}
#| eval: false
#| code-fold: false
# initialize modules
title_input <- mod_input_server("input")
# return the data and title
selected_data <- title_input$data
dataset_title <- title_input$ds_title
```

These reactive values are passed to the output modules.

## List 

`mod_list` is a simple module that collects the list of `#TidyTuesday` datasets from the `mod_input_server()`: 

### Input Controls

```{r}
#| eval: false
#| code-fold: false
# collect data
mod_list_server(id = "listviewerlite", data = selected_data)
```

### List Display

The list display comes from the [`listviewerlite` package](https://long39ng.github.io/listviewerlite/). The `mod_list_server()` accepts the `selected_data` as `data()` 

```{r}
#| eval: false
#| code-fold: false
# render data()
output$list <- renderUI({
  req(data())
  listviewerlite::listview(data())
})
```

## Table

The `mod_table` module is a Shiny module that provides interactive data table functionality for `#TidyTuesday` datasets using [the `reactable` package](https://glin.github.io/reactable/). This module provides users with a powerful interface for exploring tabular data with built-in search, sorting, and pagination capabilities.

### Input Controls

The UI is wrapped in a `bslib::card()` for consistent styling and visual organization:

```{r}
#| eval: false
#| code-fold: false
bslib::card(
  bslib::card_body(
    selectInput(
      inputId = ns("ds_input"),
      label = "Dataset",
      choices = c(c("", NULL)),
      selected = ""
    ),
    markdown("The table display below is provided by [`reactable`](https://glin.github.io/reactable/):"),
    reactable::reactableOutput(ns("table"))
  )
)
```

This creates a clean, contained interface that includes:

1. **Dataset selector dropdown** for choosing which dataset to display
2. **Informational text** crediting the `reactable` package with a helpful link
3. **Table output container** for the interactive data table

### Dataset Selector Population

The server begins by populating the dataset selector with available datasets:

```{r}
#| eval: false
#| code-fold: false
observe({
  req(data())
  ds_names <- names(data())
  updateSelectInput(
    inputId = "ds_input",
    choices = ds_names, 
    selected = ds_names[1]
  )
}) |>
  bindEvent(data())
```

This observer:

1. **Waits for data availability** using `req(data())`
2. **Extracts dataset names** from the reactive data
3. **Updates the select input** with available choices
4. **Selects the first dataset** by default
5. **Responds only to data changes** via `bindEvent()`

### Table Rendering 

The core functionality lies in the table rendering observer:

```{r}
#| eval: false
#| code-fold: false
observe({
  req(data())
  output$table <- reactable::renderReactable({
    # Table generation logic
  })
}) |>
  bindEvent(data())
```

This pattern ensures the table is re-rendered whenever the underlying data changes, but not on every reactive cycle.

### Data Processing/Validation

The table rendering includes several data processing steps:

```{r}
#| eval: false
#| code-fold: false
tryCatch({
  req(data())
  if (length(data()) == 0) {
    logr_msg(
      message = "No data available for table rendering", 
      level = "WARN")
    return(NULL)
  }
  
  ds_list <- data()
  ds_selected <- as.character(input$ds_input)
  display_ds <- ds_list[[ds_selected]]
  
  if (!is.data.frame(display_ds)) {
    logr_msg(
      message = "First dataset is not a data frame", 
      level = "ERROR")
    return(NULL)
  }
  # code continues...
})
```

This validation ensures:

1. **Data exists** before processing
2. **Dataset is properly selected** from the list
3. **Selected data is a valid data frame**
4. **Appropriate logging** for debugging and monitoring

### Performance Optimization

The module includes performance considerations for large datasets:

```{r}
#| eval: false
#| code-fold: false
display_ds <- if (nrow(display_ds) > 1000) {
  logr_msg(
    message = "Large dataset detected - limiting to first 1000 rows", 
    level = "INFO")
  head(display_ds, 1000)
} else {
  display_ds
}
```

This prevents performance issues by:

- **Limiting display to 1000 rows** for large datasets    
- **Logging the limitation** for user awareness     
- **Maintaining full functionality** for smaller datasets     

### Dynamic Subtitle 

The module creates informative subtitles based on the data context:

```{r}
#| eval: false
#| code-fold: false
subtitle <- if (length(ds_list) > 1) {
  paste("Showing:", ds_selected, "(", length(ds_list), "files total )")
} else {
  paste("File:", ds_selected)
}
```

This provides users with context about:   

- **Which dataset** they're currently viewing     
- **How many total datasets** are available     
- **Clear file identification** for reference       

### Reactable 

The module leverages `reactable`'s extensive features for an enhanced user experience:

```{r}
#| eval: false
#| code-fold: false
reactable::reactable(
  display_ds,
  searchable = TRUE,
  sortable = TRUE,
  pagination = TRUE,
  defaultPageSize = 10,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(5, 10, 20, 50),
  highlight = TRUE,
  bordered = TRUE,
  striped = TRUE,
  resizable = TRUE,
  wrap = FALSE,
  defaultColDef = reactable::colDef(
    minWidth = 100,
    headerStyle = list(background = "#f7f7f8")
  )
)
```

This configuration provides:

### Features 

1. **Search functionality** across all columns
2. **Sortable columns** for data exploration
3. **Pagination** with customizable page sizes
4. **Column resizing** for better data viewing
5. **Row highlighting** for improved readability  
6. **Striped rows** for easier scanning
7. **Bordered cells** for clear data separation
8. **Custom header styling** with subtle background color
9. **Minimum column width** to prevent cramped display
10. **No text wrapping** to maintain clean layout   
11. **Multiple page size options** (5, 10, 20, 50 rows)   
12.  **Show page size selector** for user preference    
13. **Resizable columns** for custom layouts    

### Error Handling

The UI includes comprehensive error handling:

```{r}
#| eval: false
#| code-fold: false
tryCatch({
  # UI creation code
}, error = function(e) {
  logr_msg(message = paste("Error creating table UI:", e$message), level = "ERROR")
  bslib::card(
    bslib::card_header("Table Error"),
    bslib::card_body(
      h4("Error loading table interface", class = "text-danger")
    )
  )
})
```

If UI creation fails, users see a graceful error message instead of a broken interface.

```{r}
#| eval: false
#| code-fold: false
tryCatch({
  # Table rendering logic
}, error = function(e) {
  logr_msg(
    message = paste("Error rendering table:", e$message), 
    level = "ERROR")
  reactable::reactable(
    data.frame(Error = paste("Failed to load data:", e$message)),
    searchable = FALSE,
    sortable = FALSE,
    pagination = FALSE
  )
})
```

The server error handling:

1. **Logs errors** for debugging
2. **Returns a functional table** with error information
3. **Disables interactive features** on error tables
4. **Provides user-friendly error messages**

### Integration with Main Application

In the main application, the module is integrated as:

```{r}
#| eval: false
#| code-fold: false
# In app_ui()
bslib::nav_panel("Table", mod_table_ui("table"))

# In app_server()
mod_table_server(id = "table", data = selected_data)
```

The module receives `selected_data` (a reactive containing TidyTuesday datasets) and provides a complete tabular interface within the "Table" navigation panel.

### Event-Driven Updates

The module uses `bindEvent()` to ensure efficient reactivity:

```{r}
#| eval: false
#| code-fold: false
observe({
  # Logic here
}) |>
  bindEvent(data())
```

This pattern ensures updates occur only when the underlying data changes, not on every reactive cycle.

### Requirement Validation

Strategic use of `req()` prevents processing invalid states:

```{r}
#| eval: false
#| code-fold: false
req(data())
req(input$ds_input)
```

This ensures code only executes when all necessary inputs are available.

This module demonstrates how to create robust, user-friendly data table interfaces in Shiny while maintaining clean code architecture and comprehensive error handling. The integration with `reactable` provides powerful data exploration capabilities while the modular design ensures easy integration into larger applications.

## Graphs

The `mod_plot` module provides interactive data visualization capabilities for the `#TidyTuesday` dataset list. This module is designed with a flexible interface for exploring different types of plots across multiple datasets.

### Input Controls

The UI creates three main input controls:

```{r}
#| eval: false
#| code-fold: false
selectInput(
  ns("dataset"),
  "Choose dataset:",
  choices = NULL,
  selected = NULL
)
```

This dropdown allows users to select which dataset they want to visualize. The choices are populated dynamically based on the available datasets.

```{r}
#| eval: false
#| code-fold: false
checkboxGroupInput(
  ns("plots"),
  "Select plot type:",
  choices = c(
    "types"       = "types",
    "memory"      = "mem", 
    "missing"     = "na",
    "correlation" = "cor",
    "imbalance"   = "imb",
    "numeric"     = "num",
    "categorical" = "cat"
  ),
  selected = c("types","mem","na")
)
```

This checkbox group lets users select multiple plot types simultaneously. The default selection includes three fundamental exploratory plots: column types, memory usage, and missing values analysis.

### Dynamic Plot Display

```{r}
#| eval: false
#| code-fold: false
uiOutput(ns("plots_ui"))
```

This `uiOutput` creates a dynamic container that will hold the actual plot outputs. The content is generated server-side based on user selections.

### Reactive Data Handling

The server function begins by handling both reactive and static data inputs:

```{r}
#| eval: false
#| code-fold: false
ttd_r <- if (is.reactive(ttd)) {
  logr_msg(message = "ttd is reactive", level = "DEBUG")
  ttd
} else {
  logr_msg(message = "ttd is static; wrapping in reactive()", level = "DEBUG")
  reactive(ttd)
}
```

This pattern ensures the module works whether it receives reactive data (from another module) or static data (passed directly). This flexibility is crucial for modular design.

### Dataset Selector Population

```{r}
#| eval: false
#| code-fold: false
observe({
  tryCatch({
    ds <- names(ttd_r())
    if (is.null(ds) || length(ds) == 0) {
      ds <- seq_along(ttd_r()) %>% as.character()
      logr_msg(message = "No names in ttd; using indices", level = "WARN")
    }
    updateSelectInput(
      session, "dataset",
      choices  = ds,
      selected = ds[[1]]
    )
  }, error = function(err) {
    logr_msg(message = sprintf("Error populating dataset selector: %s", err$message), level = "ERROR")
  })
}) |>
  bindEvent(ttd_r(), once = FALSE)
```

This observer:
1. **Extracts dataset names** from the reactive data
2. **Provides fallback logic** if names are missing (uses indices)
3. **Updates the select input** with available choices
4. **Selects the first dataset** by default
5. **Handles errors gracefully** with proper logging

The `bindEvent(ttd_r(), once = FALSE)` ensures this runs whenever the data changes, but not on every reactive cycle.

### Dynamic UI Generation

```{r}
#| eval: false
#| code-fold: false
output$plots_ui <- renderUI({
  tryCatch({
    req(input$dataset, input$plots)
    purrr::map(input$plots, function(plt) {
      plot_id <- ns(paste0("plt_", plt))
      tagList(
        h4(plt),
        plotOutput(plot_id, height = "300px")
      )
    })
  }, error = function(err) {
    tagList(p(style = "color:red;", "Failed to generate plot UI."))
  })
})
```

This `renderUI` creates plot containers dynamically:
1. **Validates inputs** using `req()`
2. **Maps over selected plot types** using `purrr::map()`
3. **Creates unique plot IDs** using the namespace function
4. **Generates plot containers** with headers and specified dimensions
5. **Provides error handling** with user-friendly messages

### Plot Rendering Logic

The most complex part of the module handles the actual plot generation:

```{r}
#| eval: false
#| code-fold: false
observe({
  tryCatch({
    ds_list <- req(ttd_r())
    sel_ds  <- input$dataset
    req(sel_ds %in% names(ds_list))
    df <- ds_list[[sel_ds]]
    
    purrr::walk(input$plots, function(plt) {
      out_id <- paste0("plt_", plt)
      output[[out_id]] <- renderPlot({
        tryCatch({
          inspect_plot(
            ttd = setNames(list(df), sel_ds),
            plots = plt
          )
        }, error = function(err2) {
          plot.new()
          text(0.5, 0.5, "Error generating plot", cex = 1.2, col = "red")
        })
      })
    })
  }, error = function(err) {
    logr_msg(message = sprintf("Error in plot observer: %s", err$message), level = "ERROR")
  })
}) |>
  bindEvent(list(input$dataset, input$plots), ignoreNULL = TRUE)
```

This observer:

1. **Validates the dataset selection** exists in the data   
2. **Extracts the selected dataset** as a standalone data frame   
3. **Iterates over each selected plot type** using `purrr::walk()`    
4. **Creates individual plot outputs** with unique IDs    
5. **Calls the `inspect_plot()` function** with properly formatted data   
6. **Provides plot-level error handling** with visual feedback    
7. **Responds to changes** in either dataset or plot selections   

### Integration with `inspect_plot()`

The module interfaces with the `inspect_plot()` utility function, which uses the `inspectdf` package to generate different types of exploratory plots:

- **"types"**: Column type distributions
- **"mem"**: Memory usage analysis  
- **"na"**: Missing value patterns
- **"cor"**: Correlation analysis
- **"imb"**: Feature imbalance for categorical data
- **"num"**: Numeric column summaries
- **"cat"**: Categorical column summaries

The `inspect_plot()` function handles the complexity of determining appropriate visualizations based on data characteristics and the number of datasets.

### Error Handling Strategy

The module implements comprehensive error handling at multiple levels:

1. **Input validation** using `req()` and explicit checks
2. **Graceful degradation** with fallback UI elements
3. **Detailed logging** at different severity levels
4. **User-friendly error messages** instead of technical errors
5. **Plot-level error handling** with visual placeholders

### Module Communication

The module receives data through the `ttd` parameter, which can be:   

- A reactive expression from another module   
- Static data passed directly   
- Named list of data frames (TidyTuesday format)    

This flexible input handling makes the module highly reusable across different contexts.

### Performance Considerations

The module uses several techniques to optimize performance:

- **Event-driven reactivity** with `bindEvent()` to limit unnecessary updates   
- **Lazy evaluation** with `req()` to avoid processing invalid states   
- **Efficient data handling** by extracting only the selected dataset   
- **Modular plot generation** to avoid regenerating unchanged plots   

### Usage in the Main Application

In the main app, the module is integrated as:

```{r}
#| eval: false
#| code-fold: false
# In app_ui()
bslib::nav_panel("Graphs", mod_plot_ui("viz"))

# In app_server()  
mod_plot_server("viz", ttd = selected_data)
```

The module receives `selected_data` (a reactive containing TidyTuesday datasets) and provides a complete visualization interface within the "Graphs" tab.

### Key Design Patterns

1. **Namespace Management**: Proper use of `ns()` for unique IDs
2. **Reactive Data Handling**: Flexible input processing
3. **Dynamic UI Generation**: Creating interfaces based on data
4. **Error Resilience**: Multiple layers of error handling
5. **Modular Architecture**: Clean separation of concerns
6. **Event-Driven Updates**: Efficient reactive programming

This module demonstrates sophisticated Shiny development practices while providing users with an intuitive interface for exploring their data through multiple visualization lenses.
