---
title: "Data"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

```{r}
#| label: package
#| warning: false 
#| message: false
library(ttdviewer)
library(dplyr)
library(tidyr)
library(ttmeta)
```


This document covers the `#TidyTuesday` data used in the application.

## TidyTuesday R packages 

### [tidytuesdayR](https://dslc-io.github.io/tidytuesdayR/)

The [`tidytuesdayR`](https://dslc-io.github.io/tidytuesdayR/) package is designed for, "*providing functions to quickly import data posted to the Tidy Tuesday repository.*"

```{r}
library(tidytuesdayR)
```

The `tt_load()` function downloads the data using a date input. 

```{r}
#| label: tt_load
tt_data <- tt_load("2025-07-08")
```

The downloaded object is a list, but with a few additional attributes. 

```{r}
#| label: str-tt_data
str(tt_data)
```

### Class

The `tt_data` list has it's own special class (`"tt_data"`).

```{r}
attr(tt_data, "class")
```
The `.tt` attribute prints the available datasets for the `date`.

```{r}
attr(tt_data, ".tt")
```

The `.tt` attribute also lists the available datasets, a `.files` dataset with file type and delim, the `.readme` (as html), and the `.date`:

```{r}
str(attr(tt_data, ".tt"))
```

#### .files

The `.tt` 

```{r}
attr(tt_data, ".tt") |> 
  attr(".files")
```

#### .readme

```{r}
attr(tt_data, ".tt") |> 
  attr(".readme")
```

#### .date

```{r}
attr(tt_data, ".tt") |> 
  attr(".date")
```



### [ttmeta](https://r4ds.github.io/ttmeta/)

The [`ttmeta`](https://r4ds.github.io/ttmeta/) package provides, "*a summary of each weekly TidyTuesday post, information about the articles and data sources linked in those posts, and details about the datasets themselves, including variable names and classes.*"

```{r}
library(ttmeta)
```

```{r}
#| label: get_tt_tbl
tt_meta <- get_tt_tbl(min_year = 2025L, max_year = this_year())
```

```{r}
#| label: str-tt_meta
str(tt_meta)
```

## GitHub URLs

The [#TidyTuesday repository](https://github.com/rfordatascience/tidytuesday) has [a .csv file](https://github.com/rfordatascience/tidytuesday/blob/main/static/tt_data_type.csv) with the `Week`, `Date`, `year`, `data_files`, `data_type`, and `delim`: 

This file is used to create the `tt_github_urls` data: 

```{r}
#| label: github_url
tt_github_urls <- vroom::vroom(
  file = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/static/tt_data_type.csv",
  delim = ",") |>
  # create github_url column
  dplyr::mutate(
    github_url = glue::glue(
      "https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/{year}/{Date}/{data_files}"
    )
  ) |>
  # clean names
  dplyr::rename(
    week = Week,
    date = Date
  ) 
dplyr::glimpse(tt_github_urls)
```

The `tt_github_urls` data comes from the [.csv file in TidyTuesday repo](https://github.com/rfordatascience/tidytuesday/blob/main/static/tt_data_type.csv):

### All TidyTuesday Data 

The `all_tt_data.rda` dataset contains all years (2018 - current) from the `ttmeta::get_tt_tbl()` function. This datasets contain the following columns:

```{r}
#| label: all_tt_data
dplyr::glimpse(all_tt_data)
```

### All TidyTuesday Meta 

The `all_tt_meta.rda` dataset contains all years (2018 - current) from the `ttmeta::load_tt_datasets_metadata()`. This dataset the following columns:

```{r}
#| label: all_tt_meta
dplyr::glimpse(all_tt_meta)
```

### All TidyTuesday Variable Details

The `all_tt_var_details.rda` dataset contains the following columns:

```{r}
#| label: all_tt_var_details
dplyr::glimpse(all_tt_var_details)
```

### All TidyTuesday Combined  

The `all_tt_combined.rda` dataset is `all_tt_meta` left-joined with `all_tt_data` by `year` and `week`.

```{r}
#| label: all_tt_combined
dplyr::glimpse(all_tt_combined)
```

The `clean_title` variable has been added to create an attribute we can use to join to the `all_tt_var_details` data. 

## load_tt_data()

The `load_tt_data()` function uses the `title` from `all_tt_combined`  to the return the datasets from the GitHub repo. 

### A single dataset

If the tile contains a single dataset, the list returns `ttd` with a single element: 

```{r}
#| label: ttd
ttd <- load_tt_data("Netflix Titles")
```

```{r}
#| label: ttd_nms
ttd_nms <- names(ttd)
ttd_nms
```

The `clean_title` attribute can be used to join each dataset back to `all_tt_combined` or `all_tt_data`: 

```{r}
#| label: ttd_clean_title
attr(ttd[["netflix_titles.csv"]], "clean_title")
```

### Two datasets 

If `ttd` has two datasets, we can subset the list with the name position:

```{r}
#| label: ttd2
ttd2 <- load_tt_data("Space Launches")
# store names 
ttd2_nms <- names(ttd2)
# check attr
purrr::map(.x = ttd2, .f = attr, "clean_title")
# check first dataset 
dplyr::glimpse(ttd2[[ttd2_nms[1]]])
# check second dataset
dplyr::glimpse(ttd2[[ttd2_nms[2]]])
```

### Three datasets

If there are three datasets, we can see the names are assigned to each element in the list:

```{r}
#| label: ttd3
ttd3 <- load_tt_data("Moore’s Law")
ttd3_nms <- names(ttd3)
# attr
purrr::map(.x = ttd3, .f = attr, "clean_title")
dplyr::glimpse(ttd3[[ttd3_nms[1]]])
dplyr::glimpse(ttd3[[ttd3_nms[2]]])
dplyr::glimpse(ttd3[[ttd3_nms[3]]])
```

### Many datasets

If there are more than two datasets, the same rules apply.

```{r}
#| label: ttd_many
ttd_many <- load_tt_data("LEGO database")
ttd_many_nms <- names(ttd_many)
# check attr
purrr::map(.x = ttd_many, .f = attr, "clean_title")
```

## Meta data

The `get_tt_title_meta()` function returns info on the columns in each dataset in a `ttd` list. 

```{r}
#| label: get_tt_title_meta_ttd_chr_cols
#| eval: true
ttd_meta <- get_tt_title_meta(ttd = ttd)
ttd_meta
```

```{r}
#| eval: true
ttd_meta |>
  dplyr::filter(col_type == "character") |>
  dplyr::count(clean_title, dataset, col_type, name = "chr_cols") |>
  dplyr::arrange(desc(chr_cols))
```


```{r}
#| label: get_tt_title_meta_ttd2
ttd2_meta <- get_tt_title_meta(ttd = ttd2)
head(ttd2_meta)
```


```{r}
#| label: get_tt_title_meta_ttd3
#| eval: true
ttd3_meta <- get_tt_title_meta(ttd = ttd3)
head(ttd3_meta)
```

```{r}
#| label: get_tt_title_meta_ttd_many
#| eval: true
ttd_many_meta <- get_tt_title_meta(ttd = ttd_many)
head(ttd_many_meta)
```

Below is the number of numeric columns per `dataset` and `clean_title`: 

```{r}
#| label: ttd_many_meta_num_cols
#| eval: true
ttd_many_meta |>
  dplyr::filter(col_type == "numeric") |>
  dplyr::count(clean_title, dataset, col_type, name = "num_cols") |>
  dplyr::arrange(desc(num_cols))
```


