---
title: "Data"
format: 
  html:
    toc: true
    toc-depth: 4
knitr:
  opts_chunk: 
    collapse: true
    comment: "# " 

execute:
  echo: true
  error: true

vignette: >
  %\VignetteIndexEntry{Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(tidyr)
library(inspectdf)
library(lobstr)
```

```{r}
#| label: package
library(ttdviewer)
```

The goal of `ttdviewer` is to demonstrate how to use parameterized reports (R Markdown and/or Quarto) in Shiny applications. This document covers the TidyTuesday data used in the application. '*all you need to know to download reports from your Shiny app.*'

## TidyTuesday data 

The application uses [TidyTuesday data](https://github.com/rfordatascience/tidytuesday) via the [`tidytuesdayR`](https://github.com/dslc-io/tidytuesdayR) and [`ttmeta`](https://github.com/r4ds/ttmeta) packages. 

### TidyTuesday GitHub URLs

The `tt_github_urls` data comes from the [.csv file in TidyTuesday repo](https://github.com/rfordatascience/tidytuesday/blob/main/static/tt_data_type.csv):

```{r}
#| label: tt_github_urls
dplyr::glimpse(tt_github_urls)
```

### All TidyTuesday Data 

The `all_tt_data.rda` dataset contains all years (2018 - current) from the `ttmeta::get_tt_tbl()` function. This datasets contain the following columns:

```{r}
#| label: all_tt_data
dplyr::glimpse(all_tt_data)
```

### All TidyTuesday Meta 

The `all_tt_meta.rda` dataset contains all years (2018 - current) from the `ttmeta::load_tt_datasets_metadata()`. This dataset the following columns:

```{r}
#| label: all_tt_meta
dplyr::glimpse(all_tt_meta)
```

### All TidyTuesday Variable Details

The `all_tt_var_details.rda` dataset contains the following columns:

```{r}
#| label: all_tt_var_details
dplyr::glimpse(all_tt_var_details)
```

### All TidyTuesday Combined  

The `all_tt_combined.rda` dataset is `all_tt_meta` left-joined with `all_tt_data` by `year` and `week`.

```{r}
#| label: all_tt_combined
dplyr::glimpse(all_tt_combined)
```

The `clean_title` variable has been added to create an attribute we can use to join to the `all_tt_var_details` data. 

## load_tt_data()

The `load_tt_data()` function uses the `title` from `all_tt_combined`  to the return the datasets from the GitHub repo. 

### A single dataset

If the tile contains a single dataset, the list returns `ttd` with a single element: 

```{r}
#| label: ttd
ttd <- load_tt_data("Netflix Titles")
```

```{r}
#| label: ttd_nms
ttd_nms <- names(ttd)
ttd_nms
```

The `clean_title` can be used to join back to `all_tt_combined` or `all_tt_data`: 

```{r}
#| label: ttd_clean_title
attr(ttd[["netflix_titles.csv"]], "clean_title")
```

### Two datasets 

If `ttd` has two datasets, subset the list with the name position:

```{r}
#| label: ttd2
ttd2 <- load_tt_data("Space Launches")
```

```{r}
#| label: ttd2_nms
ttd2_nms <- names(ttd2)
ttd2_nms
```

```{r}
#| label: ttd2_clean_title
purrr::map(.x = ttd2, .f = attr, "clean_title")
```

```{r}
ttd2_nms[1]
dplyr::glimpse(ttd2[[ttd2_nms[1]]])
```

```{r}
ttd2_nms[2]
dplyr::glimpse(ttd2[[ttd2_nms[2]]])
```

### Three datasets

If there are three datasets, we can see the names are assigned to each element in the list:

```{r}
#| label: ttd3
ttd3 <- load_tt_data("Moore’s Law")
```

```{r}
#| label: ttd3_nms
ttd3_nms <- names(ttd3)
ttd3_nms
```

These names can be used to subset the list:

```{r}
dplyr::glimpse(ttd3[[ttd3_nms[1]]])
```


```{r}
dplyr::glimpse(ttd3[[ttd3_nms[2]]])
```

```{r}
dplyr::glimpse(ttd3[[ttd3_nms[3]]])
```

The `clean_title` attribute will give us a joining variable:

```{r}
#| label: ttd3_clean_title
purrr::map(.x = ttd3, .f = attr, "clean_title")
```

### Many datasets

If there are more than two datasets, the same rules apply.

```{r}
#| label: ttd_many
ttd_many <- load_tt_data("LEGO database")
```

```{r}
#| label: ttd_many_nms
ttd_many_nms <- names(ttd_many)
ttd_many_nms
```

```{r}
#| label: ttd_many_nms_clean_title
purrr::map(.x = ttd_many, .f = attr, "clean_title")
```

## Meta data

The `get_tt_title_meta()` function returns info on the columns in each dataset in a `ttd` list. 

```{r}
#| label: get_tt_title_meta_ttd_chr_cols
#| eval: true
ttd_meta <- get_tt_title_meta(ttd = ttd)
ttd_meta
```

```{r}
#| eval: false
ttd_meta |>
  dplyr::filter(col_type == "character") |>
  dplyr::count(clean_title, dataset, col_type, name = "chr_cols") |>
  dplyr::arrange(desc(chr_cols))
```


```{r}
#| label: get_tt_title_meta_ttd2
ttd2_meta <- get_tt_title_meta(ttd = ttd2)
head(ttd2_meta)
```


```{r}
#| label: get_tt_title_meta_ttd3
#| eval: false
ttd3_meta <- get_tt_title_meta(ttd = ttd3)
head(ttd3_meta)
```

```{r}
#| label: get_tt_title_meta_ttd_many
#| eval: false
ttd_many_meta <- get_tt_title_meta(ttd = ttd_many)
head(ttd_many_meta)
```

Below is the number of numeric columns per `dataset` and `clean_title`: 

```{r}
#| label: ttd_many_meta_num_cols
#| eval: false
ttd_many_meta |>
  dplyr::filter(col_type == "numeric") |>
  dplyr::count(clean_title, dataset, col_type, name = "num_cols") |>
  dplyr::arrange(desc(num_cols))
```

