---
title: "Graphs"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
    fig.dpi = 320,
    fig.width = 6,
    out.width = '80%',
    fig.format =  "png",
    fig.align = 'center'
)
```


```{r}
#| label: package
#| message: false 
#| warning: false
library(ttdviewer)
library(inspectdf)
library(dplyr)
library(tidyr)
```

Below are examples of the data visualizations using `inspectdf` in the Shiny app. 

## Data 

We'll need data to demonstrate how the `inspect_plot()` function works. We'll use the `load_tt_data()` function to collect four lists of data from the [TidyTuesday](https://github.com/rfordatascience/tidytuesday) repo.

`"Netflix Titles"` has a single dataset:

```{r}
#| label: ttd_netflix_titles
#| eval: true
#| message: false 
#| warning: false
ttd <- load_tt_data("Netflix Titles")
purrr::map(.x = ttd, .f = attr, "clean_title")
# for a list with two datasets, we'll use the `"Space Launches"` data
ttd2 <- load_tt_data("Space Launches")
purrr::map(.x = ttd2, .f = attr, "clean_title")
# "Moore's Law" contains three datasets
ttd3 <- load_tt_data("Moore’s Law")
purrr::map(.x = ttd3, .f = attr, "clean_title")
# finally, we'll download the "LEGO database" data with 12 datasets into the
# ttd_many list:
ttd_many <- load_tt_data("LEGO database")
purrr::map(.x = ttd_many, .f = attr, "clean_title")
```

## Utility functions

The `inspect_plot()` function has the following utility functions: 

### Check length 

Check the length of the `ttd3` list with `ttd_length()`: 

```{r}
#| label: ttd_length
ttd_length(ttd = ttd3)
```

### Check numeric columns

Check if datasets have numerical columns: 

```{r}
#| label: check_ttd_num_cols
check_ttd_num_cols(ttd = ttd3)
```
  - If so, do they have two or more?

### Check categorical columns

Check if datasets have categorical (character) columns:

```{r}
#| label: check_ttd_cat_cols
check_ttd_cat_cols(ttd3)
```

### Check list columns

Check list columns:

```{r}
#| label: check_ttd_list_cols
check_ttd_list_cols(ttd3)
```

### Check logical columns

Check logical columns:

```{r}
#| label: check_ttd_log_cols
check_ttd_log_cols(ttd3)
```

### Check date columns

Check date columns:

```{r}
#| label: check_ttd_date_cols
check_ttd_date_cols(ttd3)
```

### check_col_types()

These are wrapped in the `check_col_types()` function:

```{r}
#| label: check_col_types_cat
identical( # check the
  x = length(check_col_types(ttd = ttd3, type = "cat")),
  y = length(check_ttd_cat_cols(ttd3))
)
```

```{r}
#| label: check_col_types_date
identical(
  x = length(check_col_types(ttd = ttd3, type = "date")),
  y = length(check_ttd_date_cols(ttd3))
)
```

```{r}
#| label: check_col_types_log
identical(
  x = length(check_col_types(ttd = ttd3, type = "log")),
  y = length(check_ttd_log_cols(ttd3))
)
```

```{r}
#| label: check_col_types_num
identical(
  x = length(check_col_types(ttd = ttd3, type = "num")),
  y = length(check_ttd_num_cols(ttd3))
)
```

The functions above allow us to collect the following: 

1. Determine the number of datasets in the `ttd` list  	
2. Determine the name of the datasets in the `ttd` list  	
3. Determine the number of numerical columns in each dataset in the `ttd` list	  
4. Determine the names of any intersecting numerical columns in each dataset in the `ttd` list	  
5. Determine the number of categorical columns in each dataset in the `ttd` list		
6. Determine the names of any intersecting categorical columns in each dataset in the `ttd` list		
7. Determine the number of logical columns in each dataset in the `ttd` list	
8. Determine the number of date columns in each dataset in the `ttd` list		
8. Determine the names of any date columns in each dataset in the `ttd` list		
9. Determine the number of list columns in each dataset in the `ttd` list	
10. Determine the names of any list columns in each dataset in the `ttd` list	

## Function logic

Using the functions above, write the custom `inspect_plot()` function and have it adhere to the following rules: 

### inspect_types()

#### `if (length(ttd) == 1)`

If one dataset in `ttd`: 

```{r}
inspectdf::inspect_types(
  df1 = ttd2[[1]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(ttd = 
    list("agencies.csv" = ttd2[[1]]), 
  plots = "types")
```


#### `if (length(ttd) == 2)`

If two datasets in `ttd`: 

  - check for intersecting columns  

  - if columns intersect, store column names as `intersecting_columns`

```{r}
if (length(intersect(names(ttd2[[1]]), names(ttd2[[2]]))) > 0) {
  intersecting_columns <- intersect(
  x = names(ttd2[[1]]), 
  y = names(ttd2[[2]])
  )
}
```

  - Then use `intersecting_columns` to subset `ttd` for the `inspect_cat()` function: 

```{r}
inspectdf::inspect_types(
  df1 = ttd2[[1]][intersecting_columns],
  df2 = ttd2[[2]][intersecting_columns],
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

#### `if (length(ttd) >= 3)`

If more than two datasets in `ttd`: 

```{r}
inspectdf::inspect_types(
  df1 = ttd3[[1]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("cpu.csv" = ttd3[[1]]),
  plots = "types"
  )
```


```{r}
inspectdf::inspect_types(
  df1 = ttd3[[2]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("gpu.csv" = ttd3[[2]]), 
  plots = "types"
  )
```


```{r}
inspectdf::inspect_types(
  df1 = ttd3[[3]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("ram.csv" = ttd3[[3]]), 
  plots = "types"
  )
```


### inspect_mem()

#### `if (length(ttd) == 1)` 

If one dataset in `ttd`:

```{r}
inspectdf::inspect_mem(
  df1 = ttd[[1]],
  df2 = NULL) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("netflix_titles.csv" = ttd[[1]]), 
  plots = "mem"
  )
```


#### `if (length(ttd) == 2)`

If two datasets in `ttd`:

```{r}
inspectdf::inspect_mem(
  df1 = ttd2[[1]], 
  df2 = ttd2[[2]]) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list(
    "agencies.csv" = ttd2[[1]],
    "launches.csv" = ttd2[[2]]
    ), 
  plots = "mem")
```


#### `if (length(ttd) >= 3)`

If more than two datasets in `ttd`:

```{r}
inspectdf::inspect_mem(
  df1 = ttd3[[1]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
```


```{r}
inspect_plot(
  ttd = list("cpu.csv" = ttd3[[1]]), 
  plots = "mem"
  )
```


```{r}
inspectdf::inspect_mem(
  df1 = ttd3[[2]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("gpu.csv" = ttd3[[2]]), 
  plots = "mem"
  )
```

```{r}
inspectdf::inspect_mem(
  df1 = ttd3[[3]],
  df2 = NULL
) |>
  inspectdf::show_plot(text_labels = TRUE)
# ... for all datasets in ttd list
```

```{r}
inspect_plot(
  ttd = list("ram.csv" = ttd3[[3]]), 
  plots = "mem"
  )
```

### inspect_cor()

#### `if (length(ttd) == 1)`

If one dataset in `ttd`:
  
  - check if there are at least 2 numeric columns in the datasets in the `ttd` list	
    - Plot the dataset with at least two numeric columns
  
```{r}
if (ttd_length(ttd2) == 2) { # check for two datasets
  # check for two numerical columns in ttd[[1]]
  if (length(check_ttd_num_cols(ttd = ttd2)[[1]]) == 2) {
    # plot first dataset
    inspectdf::inspect_cor(
      df1 = ttd2[[1]],
      df2 = NULL,
      method = "pearson",
      with_col = NULL,
      alpha = 0.05
    ) |>
      inspectdf::show_plot(text_labels = TRUE)
    # check for two numerical columns in ttd[[2]]
  } else if (length(check_ttd_num_cols(ttd = ttd2)[[2]]) >= 2) {
    # plot second dataset
    inspectdf::inspect_cor(
      df1 = ttd2[[2]],
      df2 = NULL,
      method = "pearson",
      with_col = NULL,
      alpha = 0.05
    ) |>
      inspectdf::show_plot(text_labels = TRUE)
  } else {
    logr_msg("Error", level = "INFO")
  }
}
```

```{r}
inspect_plot(
  ttd = list(
    "launches.csv" = ttd2[[2]]
    ), 
  plots = "cor"
  )
```


#### `if (length(ttd) == 2)`

If two datasets in `ttd`:

  - check if there is at least 1 numeric column in each dataset from the `ttd` list	
	
```{r}
check_col_types(ttd2, type = "num")
```

```{r}
check_col_types(ttd3, type = "num")
```

  - Run `inspectdf::inspect_cor()`if at least one numerical column is in each dataset: 

```{r}
inspectdf::inspect_cor(
  df1 = ttd3[[1]],
  df2 = ttd3[[2]],
  method = "pearson",
  with_col = NULL,
  alpha = 0.05
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list(
    "cpu.csv" = ttd3[[1]],
    "gpu.csv" = ttd3[[2]]
  ), 
  plots = "cor"
  )
```


#### `if (length(ttd) >= 3)`

If more than two datasets in `ttd`:

  - check if there is at least 1 numeric column in each of the datasets from the `ttd` list:
	
```{r}
check_col_types(ttd3, type = "num")
```

  - Run the following on all datasets with at least two numerical columns. 

```{r}
inspectdf::inspect_cor(
  df1 = ttd3[[1]],
  df2 = NULL,
  method = "pearson",
  with_col = NULL,
  alpha = 0.05
) |>
  inspectdf::show_plot(text_labels = TRUE)
```


```{r}
inspect_plot(
  ttd = list("cpu.csv" = ttd3[[1]]), 
  plots = "cor")
```


```{r}
inspectdf::inspect_cor(
  df1 = ttd3[[2]],
  df2 = NULL,
  method = "pearson",
  with_col = NULL,
  alpha = 0.05
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("gpu.csv" = ttd3[[2]]), 
  plots = "cor")
```

```{r}
inspectdf::inspect_cor(
  df1 = ttd3[[3]],
  df2 = NULL,
  method = "pearson",
  with_col = NULL,
  alpha = 0.05
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("ram.csv" = ttd3[[3]]), 
  plots = "cor")
# for all other dataset in the list
```

### inspect_imb()

#### `if (length(ttd) == 1)`

If one dataset in `ttd`:

```{r}
inspectdf::inspect_imb(
  df1 = ttd[[1]],
  df2 = NULL,
  include_na = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("netflix_titles.csv" = ttd[[1]]), 
  plots = "imb")
```


#### `if (length(ttd) == 2)`

If two datasets in `ttd`:

```{r}
inspectdf::inspect_imb(
  df1 = ttd2[[1]],
  df2 = ttd2[[2]],
  include_na = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list(
    "agencies.csv" = ttd2[[1]],
    "launches.csv" = ttd2[[2]]
    ), 
  plots = "imb")
```


#### `if (length(ttd) >= 3)`

If more than two datasets in `ttd`:

```{r}
inspectdf::inspect_imb(
  df1 = ttd3[[1]],
  df2 = NULL,
  include_na = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("cpu.csv" = ttd3[[1]]), 
  plots = "imb")
```


```{r}
inspectdf::inspect_imb(
  df1 = ttd3[[2]],
  df2 = NULL,
  include_na = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("gpu.csv" = ttd3[[2]]), 
  plots = "imb")
```

```{r}
inspectdf::inspect_imb(
  df1 = ttd3[[3]],
  df2 = NULL,
  include_na = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)

```

```{r}
inspect_plot(
  ttd = list("ram.csv" = ttd3[[3]]), 
  plots = "imb")
# ... for all datasets in ttd list
```


### inspect_num()

#### `if (length(ttd) == 1)`

If one dataset in `ttd`:

  - check if there is at least 1 numeric column in the dataset in the `ttd` list	
```{r}
check_ttd_num_cols(ttd2)
```
  
    - if so, run: 

```{r}
inspectdf::inspect_num(
  df1 = ttd2[[2]],
  df2 = NULL,
  breaks = 20,
  include_int = TRUE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list("launches.csv" = ttd2[[2]]), 
  plots = "num")
```


#### `if (length(ttd) == 2)`

If two datasets in `ttd`:

  - check if there is at least 1 numeric column in each of the two datasets in the `ttd` list	

    - if so, run: 

```{r}
inspectdf::inspect_num(
  df1 = ttd2[[1]],
  df2 = ttd2[[2]],
  breaks = 20,
  include_int = TRUE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspect_plot(
  ttd = list(
  "agencies.csv" = ttd2[[1]],
  "launches.csv" = ttd2[[2]]),
  plots = "num")
```


  - if there is only one numeric column, run the following on the dataset with the numeric column:

```{r}
inspectdf::inspect_num(
  df1 = ttd[[1]],
  df2 = NULL,
  breaks = 20,
  include_int = TRUE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

#### `if (length(ttd) >= 3)`

If more than two datasets in `ttd`:

```{r}
inspectdf::inspect_num(
  df1 = ttd3[[1]],
  df2 = NULL,
  breaks = 20,
  include_int = TRUE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspectdf::inspect_num(
  df1 = ttd3[[2]],
  df2 = NULL,
  breaks = 20,
  include_int = TRUE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

```{r}
inspectdf::inspect_num(
  df1 = ttd3[[3]],
  df2 = NULL,
  breaks = 20,
  include_int = TRUE
) |>
  inspectdf::show_plot(text_labels = TRUE)
# ... for all datasets in ttd list
```

### inspect_cat()

#### `if (length(ttd) == 1)`

If one dataset in `ttd`:

  - check if there is at least 1 categorical column in the dataset in the `ttd` list	

    - if so, run: 

```{r}
inspectdf::inspect_cat(
  df1 = ttd[[1]],
  df2 = NULL,
  include_int = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

#### `if (length(ttd) == 2)`

If two datasets in `ttd`:

  - check for intersecting columns in the two datasets in the `ttd` list	
  
```{r}
length(
  intersect(
    x = check_ttd_cat_cols(ttd2)[[1]],
    y = check_ttd_cat_cols(ttd2)[[2]]
  )
) > 0
```

  - If `TRUE`, store `intersecting_columns`  
  
```{r}
intersecting_columns <- intersect(x = names(ttd2[[1]]), y = names(ttd2[[2]]))
```

  - Use `intersecting_columns` to subset each dataset in `ttd` list.

```{r}
inspectdf::inspect_cat(
  df1 = ttd2[[1]][intersecting_columns],
  df2 = ttd2[[2]][intersecting_columns],
  include_int = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

  - if there aren't any intersecting categorical columns, run the following on the dataset with the categorical column `ttd[[?]]`:

```{r}
inspectdf::inspect_cat(
  df1 = ttd[[1]],
  df2 = NULL,
  include_int = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```

#### `if (length(ttd) >= 3)`

If more than two datasets in `ttd`:

```{r}
inspectdf::inspect_cat(
  df1 = ttd_many[[1]],
  df2 = NULL,
  include_int = FALSE
) |>
  inspectdf::show_plot(text_labels = TRUE)
```
