---
title: "Lists"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lists}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: packages
#| message: false
#| warning: false
library(ttdviewer)
library(listviewerlite)
library(lobstr)
```

This vignette covers viewing list objects in the `ttdviewer` package.

## List object

Below we'll define a list with two `#TidyTuesday` datasets and store it in  `ttd2`:

```{r}
#| label: ttd2
ttd2 <- load_tt_data("Space Launches")
```

### ctr()

`ctr()` is a custom folder tree function made with `lobstr::tree()`:

```{r}
#| label: ctr
ctr <- function(x = as.list(), depth = 10L, length = 50L) {
  lobstr::tree(x,
    max_depth = depth,
    max_length = length,
    show_environments = FALSE,
    show_attributes = FALSE,
    hide_scalar_types = TRUE,
    class_printer = crayon::blue,
    remove_newlines = TRUE,
    val_printer = crayon::green
  )
}
```

```{r}
#| label: ctr-tdd2
#| comment: ""
#| eval: false
ctr(x = ttd2)
```

```{verbatim}
<list>
├─agencies.csv: S3<spec_tbl_df/tbl_df/tbl/data.frame>
│ ├─agency<chr [74]>: "RVSN", "UNKS", "NASA", "USAF", "AE", "AFSC", "VKSR", "CALT", "FKA", "SAST", ...
│ ├─count<dbl [74]>: 1528, 904, 469, 388, 258, 247, 200, 181, 128, 105, ...
│ ├─ucode<chr [74]>: "RVSN", "GUKOS", "NASA", "USAF", "AE", "AFSC", "GUKOS", "CALT", "MOM", "SBA", ...
│ ├─state_code<chr [74]>: "SU", "SU", "US", "US", "F", "US", "RU", "CN", "RU", "CN", ...
│ ├─type<chr [74]>: "O/LA", "O/LA", "O/LA/LV/PL/S", "O/LA/S", "O/LA", "LA", "O/LA", "LA/LV/PL/E", "O/LA", "O/LA/LV/PL", ...
│ ├─class<chr [74]>: "D", "D", "C", "D", "B", "D", "D", "C", "C", "B", ...
│ ├─tstart<chr [74]>: "1960", "1986 Apr 24", "1958 Oct  1", "1947 Sep 18", "1980 Mar 26", "1961 Apr  1", "1997 Jul", "1957 Nov 11", "2004", "1993", ...
│ ├─tstop<chr [74]>: "1991 Dec", "1991", "-", "-", "*", "1992 Jul  1", "2001 Jun  1", "-", "2016 Jan  1", "-", ...
│ ├─short_name<chr [74]>: "RVSN", "UNKS", "NASA", "USAF", "Arianespace", "AFSC", "VKS RVSN", "CALT", "Roskosmos", "SAST", ...
│ ├─name<chr [74]>: "Rakentiye Vo...", "Upravleniye...", "National Aer...", "United State...", "Arianespace,...", "US Air Force...", "Voenno-Kosmi...", "Zhongguo yun...", "Federal'noe...", "Shanghai han...", ...
│ ├─location<chr [74]>: "Mosvka?", "Moskva", "Washington,...", "Washington,...", "Paris-Evry,...", "El Segundo,...", "Mosvka-Solne...", "Beijing-Nanyuan", "Moskva", "Shanghai-Min...", ...
│ ├─longitude<chr [74]>: "-", "-", "-", "-", "-", "-", "-", "-", "-", "-", ...
│ ├─latitude<chr [74]>: "-", "-", "-", "-", "-", "-", "-", "-", "-", "-", ...
│ ├─error<chr [74]>: "-", "-", "-", "-", "-", "-", "-", "-", "-", "-", ...
│ ├─parent<chr [74]>: "-", "MO", "-", "-", "-", "USAF", "RVSN", "CASC", "-", "CASC", ...
│ ├─short_english_name<chr [74]>: "-", "-", "-", "-", "Arianespace", "-", "-", "CALT", "-", "-", ...
│ ├─english_name<chr [74]>: "Strategic Ro...", "-", "-", "-", "-", "-", "Russian Mili...", "Chinese Acad...", "Roskosmos", "Shanghai Aca...", ...
│ ├─unicode_name<chr [74]>: "Ракетные вой...", "Управление н...", "National Aer...", "United State...", "Arianespace,...", "US Air Force...", "Военно-косми...", "中国运载火箭技术研究院", "Федеральное...", "上海航天技术研究院", ...
│ └─agency_type<chr [74]>: "state", "state", "state", "state", "private", "state", "state", "state", "state", "state", ...
└─launches.csv: S3<spec_tbl_df/tbl_df/tbl/data.frame>
  ├─tag<chr [5,726]>: "1967-065", "1967-080", "1967-096", "1968-042", "1968-092", "1969-062", "1970-012", "1970-070", "1971-012", "1971-054", ...
  ├─JD<dbl [5,726]>: 2439671.38, 2439725.7, 2439774.83, 2439999.69, 2440152.69, 2440425.69, 2440628.86, 2440832.86, 2440999.66, 2441111.08, ...
  ├─launch_date<dbl [5,726]>: 1967-06-29, 1967-08-23, 1967-10-11, 1968-05-23, 1968-10-23, 1969-07-23, 1970-02-11, 1970-09-03, 1971-02-17, 1971-06-08, ...
  ├─launch_year<dbl [5,726]>: 1967, 1967, 1967, 1968, 1968, 1969, 1970, 1970, 1971, 1971, ...
  ├─type<chr [5,726]>: "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", "Thor Burner 2", ...
  ├─variant<chr [5,726]>: "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", ...
  ├─mission<chr [5,726]>: "Secor Type I...", "DAPP 3419", "DAPP 4417", "DAPP 5420", "DAPP 6422", "DAPP 7421", "DAPP Block 5...", "DAPP Block 5...", "DAPP Block 5...", "P70-1", ...
  ├─agency<chr [5,726]>: "US", "US", "US", "US", "US", "US", "US", "US", "US", "US", ...
  ├─state_code<chr [5,726]>: "US", "US", "US", "US", "US", "US", "US", "US", "US", "US", ...
  ├─category<chr [5,726]>: "O", "O", "O", "O", "O", "O", "O", "O", "O", "O", ...
  └─agency_type<chr [5,726]>: "state", "state", "state", "state", "state", "state", "state", "state", "state", "state", ...
```

### listviewerlite::listview()

The `listview()` from `listviewerlite` provides an HTML collapsible object:

```{r}
#| label: listviewerlite--listview-tdd2
listviewerlite::listview(x = ttd2)
```

### Three datasets

Below we create a list with three `#TidyTuesday` datasets and store it in  `ttd3`:

```{r}
#| label: ttd3
ttd3 <- load_tt_data("Moore’s Law")
```

```{r}
#| label: ctr-tdd3
#| comment: ""
#| eval: false
ctr(x = ttd3)
```

```{verbatim}
<list>
├─cpu.csv: S3<spec_tbl_df/tbl_df/tbl/data.frame>
│ ├─processor<chr [176]>: "MP944 (20-bi...", "Intel 4004 (...", "Intel 8008 (...", "NEC μCOM-4 (...", "Toshiba TLCS...", "Intel 4040 (...", "Motorola 680...", "Intel 8080 (...", "TMS 1000 (4-...", "MOS Technolo...", ...
│ ├─transistor_count<dbl [176]>: NA, 2250, 3500, 2500, 2800, 3000, 4100, 6000, 8000, 4528, ...
│ ├─date_of_introduction<dbl [176]>: 1970, 1971, 1972, 1973, 1973, 1974, 1974, 1974, 1974, 1975, ...
│ ├─designer<chr [176]>: "Garrett AiRe...", "Intel", "Intel", "NEC", "Toshiba", "Intel", "Motorola", "Intel", "Texas Instru...", "MOS Technology", ...
│ ├─process<dbl [176]>: NA, 10000, 10000, 7500, 6000, 10000, 6000, 6000, 8000, 8000, ...
│ └─area<dbl [176]>: NA, 12, 14, NA, 32, 12, 16, 20, 11, 21, ...
├─gpu.csv: S3<spec_tbl_df/tbl_df/tbl/data.frame>
│ ├─processor<chr [112]>: "µPD7220 GDC", "ARTC HD63484", "YM7101 VDP", "Tom & Jerry", "VDP1", "Sony GPU", "NV1", "Reality Copr...", "PowerVR", "Voodoo Graphics", ...
│ ├─transistor_count<dbl [112]>: 40000, 60000, 1e+05, 750000, 1e+06, 1e+06, 1e+06, 2600000, 1200000, 1e+06, ...
│ ├─date_of_introduction<dbl [112]>: 1982, 1984, 1988, 1993, 1994, 1994, 1995, 1996, 1996, 1996, ...
│ ├─designer_s<chr [112]>: "NEC", "Hitachi", "Sega", "Flare", "Sega", "Toshiba", "Nvidia, Sega", "SGI", "VideoLogic", "3dfx", ...
│ ├─manufacturer_s<chr [112]>: "NEC", "Hitachi", "Yamaha", "IBM", "Hitachi", "LSI", "SGS", "NEC", "NEC", "TSMC", ...
│ ├─process<dbl [112]>: 5000, NA, NA, NA, 500, 500, 500, 350, 350, 500, ...
│ ├─area<dbl [112]>: NA, NA, NA, NA, NA, NA, 90, 81, NA, NA, ...
│ └─ref<chr [112]>: "[107]", "[108]", "[109]", "[109]", "[110][111]", "[112][113][114]", "[110]", "[115]", "[116]", "[117][118]", ...
└─ram.csv: S3<spec_tbl_df/tbl_df/tbl/data.frame>
  ├─chip_name<chr [47]>: "N/A", "N/A", "?", "SP95", "TMC3162", "?", "?", "?", "?", "1101", ...
  ├─capacity_bits<dbl [47]>: 1, 1, 8, 16, 16, NA, 256, 64, 144, 256, ...
  ├─bit_units<chr [47]>: "Bits", "Bits", "Bits", "Bits", "Bits", "NA", "Bits", "Bits", "Bits", "Bits", ...
  ├─ram_type<chr [47]>: "SRAM (cell)", "DRAM (cell)", "SRAM (bipolar)", "SRAM (bipolar)", "SRAM (TTL)", "SRAM (MOS)", "DRAM (IC)", "SRAM (PMOS)", "SRAM (NMOS)", "SRAM (PMOS)", ...
  ├─transistor_count<dbl [47]>: 6, 1, 48, 80, 96, NA, 256, 384, 864, 1536, ...
  ├─date_of_introduction<dbl [47]>: 1963, 1965, 1965, 1965, 1966, 1966, 1968, 1968, 1968, 1969, ...
  ├─manufacturer_s<chr [47]>: "Fairchild", "Toshiba", "SDS, Signetics", "IBM", "Transitron", "NEC", "Fairchild", "Fairchild", "NEC", "Intel", ...
  ├─process<dbl [47]>: NA, NA, NA, NA, NA, NA, NA, NA, NA, 12000, ...
  ├─area<dbl [47]>: NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ...
  └─ref<chr [47]>: "[162]", "[163][164]", "[162]", "[165]", "[160]", "[159]", "[160]", "[159]", "[159]", "[166][167][168]", ...
```

```{r}
#| label: listviewerlite-listview-tdd3
listviewerlite::listview(ttd3)
```
