---
title: "Lists"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lists}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: packages
#| message: false
#| warning: false
library(ttdviewer)
library(listviewerlite)
library(lobstr)
```

This vignette covers viewing list objects in the `ttdviewer` package.

## List object

Below we'll define a list with two `#TidyTuesday` datasets and store it in  `ttd2`:

```{r}
#| label: ttd2
ttd2 <- load_tt_data("Space Launches")
```

### ctr()

`ctr()` is a custom folder tree function made with `lobstr::tree()`:

```{r}
#| label: ctr
ctr <- function(x = as.list(), depth = 10L, length = 50L) {
  lobstr::tree(x,
    max_depth = depth,
    max_length = length,
    show_environments = FALSE,
    show_attributes = FALSE,
    hide_scalar_types = TRUE,
    class_printer = crayon::blue,
    remove_newlines = TRUE,
    val_printer = crayon::green
  )
}
```

```{r}
#| label: ctr-tdd2
#| comment: ""
ctr(x = ttd2)
```

### listviewerlite::listview()

The `listview()` from `listviewerlite` provides an HTML collapsible object:

```{r}
#| label: listviewerlite--listview-tdd2
listviewerlite::listview(x = ttd2)
```

### Three datasets

Below we create a list with three `#TidyTuesday` datasets and store it in  `ttd3`:

```{r}
#| label: ttd3
ttd3 <- load_tt_data("Moore’s Law")
```

```{r}
#| label: ctr-tdd3
#| comment: ""
ctr(x = ttd3)
```

```{r}
#| label: listviewerlite-listview-tdd3
listviewerlite::listview(ttd3)
```
