---
title: "Reports"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Reports}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 75
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
#| label: package
#| message: false 
#| warning: false
library(dplyr)
library(tidyr)
library(inspectdf)
library(lobstr)
library(ttdviewer)
```

The report generation in the #TidyTuesday data viewer app consists of three
interconnected modules that work together to provide users with
customizable, downloadable reports. This method demonstrates patterns for
format selection, system capability detection, and reliable file generation
with comprehensive error handling.

## Module Architecture Overview

The report generation system consists of three main components:

1.  **`mod_report_input`**: Format selection interface\
2.  **`mod_report_desc`**: Dynamic format description and system status\
3.  **`mod_report_download`**: Report generation and download functionality

```{r}
#| label: module-structure
#| message: false 
#| warning: false
#| echo: false 
#| eval: true
#| comment: ""
lobstr::ast(
  launch(
    app_ui(
      mod_report_input_ui(),
      mod_report_desc_ui(),
      mod_report_download_ui()
    ),
    app_server(
      mod_report_input_server(),
      mod_report_desc_server(),
      mod_report_download_server()
    )
  )
)
```

These three modules work together with a sophisticated backend utility
logic that handles multiple rendering formats, quarto system detection, and
graceful error handling.

## mod_report_input - the format selector

The `mod_report_input` handles the format selection input from the user.

### UI

The input module provides a `selectInput()` interface for format selection:

```{r}
#| label: mod_report_input_ui
selectInput(
  inputId = ns("format"),
  label = strong("Download Report:"),
  choices = list(
    "R Markdown" = "rmarkdown",
    "Quarto" = "quarto"
  ),
  selected = "rmarkdown",
  width = "100%"
)
```

The input includes bold, User-friendly labels to emphasizes the
functionality. The default selected input (R Markdown) maximizes
compatibility.

### Server

The server demonstrates a simple, clean reactive pattern:

```{r}
#| label: mod_report_input_server
mod_report_input_server <- function(id) {
  moduleServer(id, function(input, output, session) {
    return(
      reactive(
        list("format" = input$format)
      )
    )
  })
}
```

The `input$format` is returned directly via a named list (instead of
internal state management). Debugging support is provided with
comprehensive logging.

## mod_report_desc - description generator

```{r}
#| label: mod_report_desc-structure
#| message: false 
#| warning: false
#| echo: false 
#| eval: true
#| comment: ""
lobstr::ast(
  launch(
    app_ui(
      mod_report_input_ui(),
      mod_report_desc_ui()
    ),
    app_server(
      mod_report_input_server(),
      mod_report_desc_server(
        quarto_available()
      )
    )
  )
)
```

### UI

The description module uses complete dynamic UI generation:

```{r}
#| label: mod_report_desc_ui
mod_report_desc_ui <- function(id) {
  ns <- NS(id)
  tagList(
    uiOutput(outputId = ns("desc"))
  )
}
```

This minimal UI delegates all content generation to the server, enabling
complete dynamic control over presentation.

### Server

The server contains sophisticated logic for system detection and adaptive
messaging. The reactive expression orchestrates multiple outputs:

1.  Checks the system capability detection through `quarto_available()`\
2.  Format-specific messaging based on selection\
3.  Visual styling based on system state

```{r}
#| label: desc_inputs
desc_inputs <- reactive({
  req(format())
  # System capability detection
  # Format-specific descriptions  
  # Alert styling logic
  return(list(
    "quarto_status" = quarto_status,
    "description" = description,
    "alert_class" = alert_class
  ))
})
```

After `quarto_status` is determined by the `quarto_available()` function,
the value is stored as `available` with a `message`:

```{r}
#| label: quarto_status
quarto_status <- if (format()$format == "quarto") {
  if (quarto_available()) {
    list(
      available = TRUE,
      message = "Quarto is available and ready to use."
    )
  } else {
    list(
      available = FALSE,
      message = "Quarto not detected. Will fall back to R Markdown if selected."
    )
  }
} else {
  list(available = TRUE, message = "Quarto is available and ready to use.")
}
```

The conditional checking behavior identifies when Quarto is selected and
provides graceful fallback messaging when it's unavailable.

The `description` returns information for `rmarkdown` and `quarto` with
bootstrap `icon` for visual identification, `title`, informative
`desc`riptions to guide user choice, and context-aware `note` that reflect
system state.

```{r}
#| label: description
description <- switch(format()$format,
  "rmarkdown" = list(
    icon = "bi-file-earmark-code",
    title = "R Markdown",
    desc = "Traditional R Markdown report. Compatible with all R installations.",
    note = "Recommended for maximum compatibility."
  ),
  "quarto" = list(
    icon = "bi-file-earmark-richtext", 
    title = "Quarto",
    desc = "Modern scientific publishing with enhanced features and interactivity.",
    note = quarto_status$message
  )
)
```

In the UI, the observer creates `div()`s with Bootstrap classes for
styling, layout, and typography. The `bindEvent()` provides event-driven
updates for optimal performance.

```{r}
#| label: output-desc
observe({
  output$desc <- renderUI({
    tags$div(
      class = paste(
        "alert border-start border-primary border-3 py-2",
        desc_inputs()$alert_class
      ),
      tags$div(
        class = "d-flex align-items-center mb-1",
        tags$i(class = paste("bi", desc_inputs()$description$icon, "me-2")),
        tags$strong(desc_inputs()$description$title)
      ),
      tags$p(desc_inputs()$description$desc, class = "mb-1 small"),
      tags$small(desc_inputs()$description$note, class = "text-muted")
    )
  })
}) |>
  bindEvent(format())
```

## mod_report_download - the rendering engine

```{r}
#| label: mod_report_download-structure
#| message: false 
#| warning: false
#| echo: false 
#| eval: true
#| comment: ""
lobstr::ast(
  launch(
    app_ui(
      mod_report_input_ui(),
      mod_report_download_ui()
    ),
    app_server(
      mod_report_input_server(),
      mod_report_download_server(
        quarto_available(),
        inspect_plot(),
        render_report(
          get_template_path(),
          create_fallback_template(),
          validate_and_clean_params(),
          render_rmarkdown_report(),
          render_quarto_report(),
          create_error_report()
        ),
        create_error_report()
      )
    )
  )
)
```

### UI

The download module provides an input with `downloadButton()`, some custom
`div()` classes, and informative help text describing report contents:

```{r}
#| label: mod_report_download_ui
mod_report_download_ui <- function(id) {
  ns <- NS(id)
  tagList(
    tags$div(
      class = "d-grid",
      downloadButton(
        ns("download_report"),
        "Download Report",
        class = "btn btn-primary",
        icon = icon("download")
      )
    ),
    tags$small(
      class = "text-muted",
      "Report includes: data visualization, summary statistics, data preview, and variable information."
    )
  )
}
```

### Server

The download handler (`downloadHandler()`) can be paired with any download
UI function (`downloadButton()` or `downloadLink()`). Every
`downloadHandler()` includes `filename` and `content` functions.

-   The `filename` portion handles file generation and error handling. The
    `filename` function generates a data-driven file name based on dataset
    content, timestamp inclusion for version control, safe character
    handling for cross-platform compatibility, and a graceful fallback for
    error conditions.

```{r}
#| label: filename
filename = function() {
  tryCatch({
    current_data <- data()
    actual_title <- if (!is.null(current_data) && length(current_data) > 0) {
      first_dataset_name <- names(current_data)[1]
      if (!is.null(first_dataset_name) && first_dataset_name != "") {
        gsub("\\.csv$", "", first_dataset_name)
      } else {
        "tidytuesday_dataset"
      }
    } else {
      "tidytuesday_dataset"
    }
    
    clean_title <- gsub("[^A-Za-z0-9_-]", "_", actual_title)
    timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
    format_suffix <- switch(input$report_format,
      "rmarkdown" = "rmd",
      "quarto" = "qto"
    )
    
    filename <- paste0(clean_title, "_", format_suffix, "_", timestamp, ".html")
    return(filename)
  }, error = function(e) {
    return("tidytuesday_report.html")
  })
}
```

-   The `content` portion includes data validation before processing (with
    format detection and automatic fallback). The structured `params` list
    are passed to each template and include `dataset_title`, `title`,
    `data_list`, `plots`, and `plot_type`:

```{r}
#| label: content
content = function(file) {
  tryCatch({
    # data validation
    current_data <- data()
    if (is.null(current_data) || length(current_data) == 0) {
      stop("No data available. Please select a dataset first.")
    }
    
    # format detection
    if (input$report_format == "quarto") {
      if (!quarto_available()) {
        # fallback
        actual_format <- "rmarkdown"
        showNotification("Quarto not available. Generating R Markdown report instead.", 
                        type = "warning", duration = 5)
      } else {
        actual_format <- "quarto"
      }
    } else {
      actual_format <- input$report_format
    }
    
    # progress notification
    showNotification(paste("Generating", tools::toTitleCase(actual_format), "report..."),
                    type = "message", duration = 8, id = "report_progress")
    
    # parameter preparation
    params <- list(
      dataset_title = actual_dataset_title,
      title = paste("TidyTuesday Report:", actual_dataset_title),
      data_list = current_data,
      plots = plots,
      plot_type = current_plot_type
    )
    
    # render 
    tryCatch({
      render_report(format = actual_format, output_file = file, params = params)
    }, error = function(render_error) {
      if (actual_format == "quarto") {
        render_report(format = "rmarkdown", output_file = file, params = params)
        showNotification("Quarto failed, generated R Markdown report instead.", 
                        type = "warning", duration = 5)
      } else {
        stop(render_error)
      }
    })
  }, error = function(e) {
    # fallback
    create_error_report(file, e$message, actual_dataset_title %||% "Unknown", input$report_format)
  })
}
```

The progress notifications provide real-time user feedback with multi-level
error handling, graceful recovery, and error report generation when all
else fails.

## Rending Reports

The core rendering function coordinates multiple rendering strategies:

1.  The `template_path` is passed to `get_template_path()` (either R
    Markdown or Quarto) and, if a file doesn't exist, the template
    resolution is handled with `create_fallback_template()`.\
2.  The `params` are passed to `validate_and_clean_params()` to ensure the
    correct values are passed into the report.\
3.  Conditional logic handles the specific `format` selected (either
    `render_rmarkdown_report()` or `render_quarto_report()`)

```{r}
#| label: render_report
render_report <- function(format = c("rmarkdown", "quarto"), output_file, params = list(), template_path = NULL) {
  format <- match.arg(format)
  
  # resolution
  if (is.null(template_path)) {
    template_path <- get_template_path(format)
  }
  
  # validation with fallback 
  if (!file.exists(template_path)) {
    template_path <- create_fallback_template(format)
  }
  
  # params validation and cleaning
  params <- validate_and_clean_params(params)
  
  # format-specific rendering
  if (format == "rmarkdown") {
    render_rmarkdown_report(template_path, output_file, params)
  } else {
    render_quarto_report(template_path, output_file, params)
  }
}
```

### R Markdown

R Markdown rendering is handled by render_rmarkdown_report`()`, which
creates a clean environment for rendering and adds any required libraries:

```{r}
#| label: render_rmarkdown_report
render_rmarkdown_report <- function(template_path, output_file, params) {
  
  logr_msg(
    message = "Rendering R Markdown report", 
    level = "DEBUG")
  
  logr_msg(
    message = paste("Output file:", output_file), 
    level = "DEBUG")

  # clean environment
  render_env <- new.env(parent = globalenv())

  # add libs/reqs
  render_env$library <- library
  render_env$require <- require

  # log parameter details
  logr_msg(
    message = paste("Rendering with dataset_title:", params$dataset_title), 
    level = "DEBUG")
  logr_msg(
    message = paste("Rendering with title:", params$title), 
    level = "DEBUG")

  rmarkdown::render(
    input = template_path,
    output_file = output_file,
    params = params,
    envir = render_env,
    quiet = FALSE, 
    clean = TRUE
  )
}
```

### Quarto

Quarto rendering is handled by `render_quarto_report()` and has a second
method as a fallback strategy (`render_quarto_system_call()`):

```{r}
#| label: render_quarto_report
render_quarto_report <- function(template_path, output_file, params) {
  # method 1 using the quarto R package
  if (requireNamespace("quarto", quietly = TRUE)) {
    quarto::quarto_render(
      input = temp_qmd,
      output_file = basename(output_file),
      execute_params = params,
      quiet = TRUE
    )
  } else {
    # method 2 system call fallback
    render_quarto_system_call(template_path, output_file, params)
  }
}
```

**Multiple Rendering Strategies:**

-   **R package method** for integrated environments

-   **System call fallback** for command-line installations

-   **Alternative rendering** using R Markdown compatibility

-   **Comprehensive error handling** with automatic fallbacks

### Quarto Detection

The `quarto_available()` function implements comprehensive system
detection:

```{r}
#| label: quarto_available
quarto_available <- function() {
  # 1. Check for quarto R package
  if (requireNamespace("quarto", quietly = TRUE)) {
    tryCatch({
      quarto::quarto_version()
      return(TRUE)
    }, error = function(e) {
      # Package exists but not functional
    })
  }
  
  # 2. Check system PATH
  result <- system2("quarto", args = "--version", stdout = TRUE, stderr = TRUE, timeout = 5)
  if (!is.null(result) && length(result) > 0) {
    return(TRUE)
  }
  
  # 3. Check common installation paths
  quarto_paths <- c(
    "/usr/local/bin/quarto",
    "/usr/bin/quarto",
    "C:/Program Files/Quarto/bin/quarto.exe"
  )
  
  for (path in quarto_paths) {
    if (file.exists(path)) {
      return(TRUE)
    }
  }
  
  return(FALSE)
}
```

**Multi-Level Detection:**

-   **R package detection** for integrated environments

-   **System PATH checking** with timeout protection

-   **Common path verification** for standard installations

-   **Graceful failure** with detailed logging

### Error Handling and Report 

When all rendering attempts fail, the system generates an informative error
report:

```{r}
#| label: create_error_report
create_error_report <- function(file, error_msg, dataset_title = "Unknown", format = "rmarkdown") {
  error_html <- paste0(
    '<!DOCTYPE html>
    <html>
    <head>
      <title>Report Generation Error</title>
      <link href="https://cdn.jsdelivr.net/npm/bootstrap@5.1.3/dist/css/bootstrap.min.css" rel="stylesheet">
    </head>
    <body>
      <div class="error-container">
        <h1 class="text-danger">Report Generation Failed</h1>
        <div class="alert alert-danger" role="alert">
          <h4 class="alert-heading">Error Details</h4>
          <p class="mb-0">', error_msg, '</p>
        </div>
        <div class="alert alert-info mt-4" role="alert">
          <h6 class="alert-heading">Troubleshooting Tips</h6>
          <ul class="mb-0">
            <li>Try selecting a different dataset</li>
            <li>Try a different report format</li>
            <li>Check your internet connection</li>
            <li>Contact support if the problem persists</li>
          </ul>
        </div>
      </div>
    </body>
    </html>'
  )
  
  writeLines(error_html, file)
}
```

**User-Friendly Error Handling:**

-   **Professional HTML presentation** with Bootstrap styling

-   **Clear error information** without technical jargon

-   **Actionable troubleshooting tips** for users

-   **Consistent visual design** matching the application

### Fallback Templates

The system includes automatic template generation when templates are
missing:

```{r}
#| label: create_fallback_template
create_fallback_template <- function(format = c("rmarkdown", "quarto")) {
  format <- match.arg(format)
  
  if (format == "rmarkdown") {
    return(create_fallback_rmd_template())
  } else {
    return(create_fallback_qmd_template())
  }
}
```

**Resilient Template System:**

-   **Automatic template generation** when missing

-   **Format-specific templates** with appropriate syntax

-   **Minimal viable reports** that always work

-   **Graceful degradation** from rich to basic functionality

## Main Application Integration

The modules are integrated in the `app_ui()` and `app_server()` below:

```{r}
#| label: in-app_ui
# in app_ui()
bslib::sidebar(
  mod_input_ui("input"),
  mod_report_input_ui("rep_form"),
  br(),
  mod_report_desc_ui("rep_desc"),
  mod_report_download_ui("rep_dwnld")
)

# in app_server()
report_format <- mod_report_input_server("rep_form")

mod_report_desc_server("rep_desc", format = report_format)

mod_report_download_server("rep_dwnld", 
  format = report_format,
  data = selected_data,
  selected_plot_type = viz_result,
  dataset_title = dataset_title
)
```

**Clean Integration Patterns:**

-   **Sidebar placement** for persistent access

-   **Reactive data flow** between modules

-   **Logical module ordering** for user workflow

-   **Comprehensive parameter passing** for functionality

## Features

+-------------------+------------------------------------+----------------+
| Feature           | Implementation                     | Goal           |
+===================+====================================+================+
| **Enhancements**  | -   Basic functionality works in   | Ensures        |
|                   |     all environments               | universal      |
|                   |                                    | compatibility  |
|                   | -   Advanced features available    | while          |
|                   |     when systems support them      | leveraging     |
|                   |                                    | advanced       |
|                   | -   Graceful fallbacks maintain    | capabilities   |
|                   |     functionality                  | when available |
+-------------------+------------------------------------+----------------+
| **Caching**       | -   System detection results are   | Optimizes      |
|                   |     cached during app sessions     | performance by |
|                   |                                    | avoiding       |
|                   | -   Template resolution cached for | redundant      |
|                   |     performance                    | operations and |
|                   |                                    | system calls   |
|                   | -   Parameter validation cached    |                |
|                   |     for repeated use               |                |
+-------------------+------------------------------------+----------------+
| **Logging**       | -   `DEBUG` level for development  | Provide        |
|                   |     and troubleshooting            | detailed       |
|                   |                                    | traceability   |
|                   | -   `INFO` level for normal        | for debugging, |
|                   |     operation tracking             | monitoring,    |
|                   |                                    | and user       |
|                   | -   `WARN` level for recoverable   | feedback       |
|                   |     issues                         |                |
|                   |                                    |                |
|                   | -   `ERROR` level for serious      |                |
|                   |     problems                       |                |
|                   |                                    |                |
|                   | -   `SUCCESS` level for            |                |
|                   |     confirmation of completion     |                |
+-------------------+------------------------------------+----------------+
| **User            | -   Real-time feedback via         | Maintain user  |
| Experience**      |     notifications                  | engagement and |
|                   |                                    | provide clear  |
|                   | -   Clear progress indication      | guidance       |
|                   |     during generation              | throughout the |
|                   |                                    | report         |
|                   | -   Informative error messages     | generation     |
|                   |     with actionable advice         | workflow       |
|                   |                                    |                |
|                   | -   Consistent visual design       |                |
|                   |     throughout the process         |                |
+-------------------+------------------------------------+----------------+
| **Lazy Loading**  | -   System detection only occurs   | Minimize       |
|                   |     when needed                    | resource usage |
|                   |                                    | and improve    |
|                   | -   Template loading deferred      | application    |
|                   |     until generation               | startup time   |
|                   |                                    |                |
|                   | -   Parameter validation cached    |                |
|                   |     for reuse                      |                |
+-------------------+------------------------------------+----------------+
| **Rendering**     | -   Clean environments for         | Prevent memory |
|                   |     rendering isolation            | leaks and      |
|                   |                                    | ensure stable  |
|                   | -   Resource cleanup after         | performance    |
|                   |     generation                     | across         |
|                   |                                    | multiple       |
|                   | -   Memory management for large    | report         |
|                   |     datasets                       | generations    |
+-------------------+------------------------------------+----------------+
| **Timeout         | -   System calls protected with    | Prevent        |
| Management**      |     timeouts                       | application    |
|                   |                                    | hanging and    |
|                   | -   Rendering processes with       | provide        |
|                   |     reasonable limits              | responsive     |
|                   |                                    | user           |
|                   | -   User feedback for long-running | experience     |
|                   |     operations                     |                |
+-------------------+------------------------------------+----------------+
| **Input           | -   Parameter sanitization for all | Ensure         |
| Validation**      |     user inputs                    | security and   |
|                   |                                    | prevent        |
|                   | -   Filename cleaning for safe     | malicious      |
|                   |     file operations                | input from     |
|                   |                                    | compromising   |
|                   | -   Path validation for template   | the system     |
|                   |     access                         |                |
+-------------------+------------------------------------+----------------+
| **Environment     | -   Clean rendering environments   | Maintain       |
| Isolation**       |     for security                   | security       |
|                   |                                    | boundaries and |
|                   | -   Restricted system access       | prevent        |
|                   |     during generation              | interference   |
|                   |                                    | between report |
|                   | -   Temporary file management with | generations    |
|                   |     cleanup                        |                |
+-------------------+------------------------------------+----------------+

---------------------------------------------------------------------------

## Utility functions

The `get_template_path()` function return the specified

```{r}
#| label: get_template_path_rmd
identical(
  # response from function
  x = get_template_path("rmarkdown"),
  # response from system.file()
  y = system.file("rmarkdown", "report_template.Rmd",
    package = "ttdviewer"
  )
)
```

```{r}
#| label: get_template_path_qmd
identical(
  # response from function
  x = get_template_path("quarto"),
  # response from system.file()
  y = system.file("quarto", "report_template.qmd",
    package = "ttdviewer"
  )
)
```

```{r}
#| label: rmd_create_fallback_template
rmd_temp <- create_fallback_template(format = "rmarkdown")
basename(rmd_temp)
```

```{r}
#| label: qmd_create_fallback_template
qmd_temp <- create_fallback_template(format = "quarto")
basename(qmd_temp)
```

```{r}
#| label: unlink_create_fallback_template
#| echo: false
#| results: hide
unlink(rmd_temp)
unlink(qmd_temp)
```
